State of the Practice for Medical Imaging Software

### 2022-03-23 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Updated captions for figures where the order of the entries needed to be clarified.  Closes #115

Table \ref{tab_final_list} shows the 29 software packages that we measured,
along with summary data collected in the year 2020. We arrange the items in
descending order of LOC. We found the initial release dates (Rlsd) for most
projects and marked the two unknown dates with ``?''. We used the date of the
latest change to each code repository to decide the latest update. We found
funding information (Fnd) for only eight projects.  For the number of
contributors (NOC) we considered anyone who made at least one accepted commit as
a contributor. The NOC is not usually the same as the number of long-term
project members, since many projects received change requests and code from the
community.  For the supported OS, 25 packages work on all three OSs: Windows
(W), macOS (M), and Linux (L). Although the usual approach to cross-platform
compatibility was to work natively on multiple OSes, five projects achieved
platform-independence via web applications. The full measurement data for all
packages is available in \citet{Dong2021-Data}.

\begin{table}[!ht]
\centering
\begin{tabular}{p{6cm}lllllllll}
\hline
\multirow{2}{*}{Software} & \multirow{2}{*}{Rlsd} & \multirow{2}{*}{Updated} & \multirow{2}{*}{Fnd} & \multirow{2}{*}{NOC} & \multirow{2}{*}{LOC} & \multicolumn{3}{c}{OS} & \multirow{2}{*}{Web} \\ \cline{7-9}
 &  &  &  &  &  & W & M & L &  \\ \hline
ParaView \citep{Ahrens2005} & 2002 & 2020-10 & X & 100 & 886326 & X & X & X & X \\
Gwyddion \citep{Nevcas2012} & 2004 & 2020-11 &  & 38 & 643427 & X & X & X &  \\
Horos \citep{horosproject2020} & ? & 2020-04 &  & 21 & 561617 &  & X &  &  \\
OsiriX Lite \citep{PixmeoSARL2019} & 2004 & 2019-11 &  & 9 & 544304 &  & X &  &  \\
3D Slicer \citep{Kikinis2014} & 1998 & 2020-08 & X & 100 & 501451 & X & X & X &  \\
Drishti \citep{Limaye2012} & 2012 & 2020-08 &  & 1 & 268168 & X & X & X &  \\
Ginkgo CADx \citep{Wollny2020} & 2010 & 2019-05 &  & 3 & 257144 & X & X & X &  \\
GATE \citep{Jan2004} & 2011 & 2020-10 &  & 45 & 207122 &  & X & X &  \\
3DimViewer \citep{TESCAN2020} & ? & 2020-03 & X & 3 & 178065 & X & X &  &  \\
medInria \citep{Fillard2012} & 2009 & 2020-11 &  & 21 & 148924 & X & X & X &  \\
BioImage Suite Web \citep{Papademetris2005} & 2018 & 2020-10 & X & 13 & 139699 &
X & X & X & X \\
Weasis \citep{Roduit2021} & 2010 & 2020-08 &  & 8 & 123272 & X & X & X &  \\
AMIDE \citep{Loening2017} & 2006 & 2017-01 &  & 4 & 102827 & X & X & X &  \\
XMedCon \citep{Nolf2003} & 2000 & 2020-08 &  & 2 & 96767 & X & X & X &  \\
ITK-SNAP \citep{Yushkevich2006} & 2006 & 2020-06 & X & 13 & 88530 & X & X & X &  \\
Papaya \citep{UTHSCSA2019} & 2012 & 2019-05 &  & 9 & 71831 & X & X & X &  \\
OHIF Viewer \citep{Ziegler2020} & 2015 & 2020-10 &  & 76 & 63951 & X & X & X & X \\
SMILI \citep{Chandra2018} & 2014 & 2020-06 &  & 9 & 62626 & X & X & X &  \\
INVESALIUS 3 \citep{Amorim2015} & 2009 & 2020-09 &  & 10 & 48605 & X & X & X &  \\
dwv \citep{Martelli2021} & 2012 & 2020-09 &  & 22 & 47815 & X & X & X & X \\
DICOM Viewer \citep{Afsar2021} & 2018 & 2020-04 & X & 5 & 30761 & X & X & X &  \\
MicroView \citep{ParallaxInnovations2020} & 2015 & 2020-08 &  & 2 & 27470 & X & X & X &  \\
MatrixUser \citep{Liu2016} & 2013 & 2018-07 &  & 1 & 23121 & X & X & X &  \\
Slice:Drop \citep{Haehn2013} & 2012 & 2020-04 &  & 3 & 19020 & X & X & X & X \\
dicompyler \citep{Panchal2010} & 2009 & 2020-01 &  & 2 & 15941 & X & X &  &  \\
Fiji \citep{Schindelin2012} & 2011 & 2020-08 & X & 55 & 10833 & X & X & X &  \\
ImageJ \citep{Rueden2017} & 1997 & 2020-08 & X & 18 & 9681 & X & X & X &  \\
MRIcroGL \citep{Rorden2021} & 2015 & 2020-08 &  & 2 & 8493 & X & X & X &  \\
DicomBrowser \citep{Archie2012} & 2012 & 2020-08 &  & 3 & 5505 & X & X & X &  \\ \hline
\end{tabular}
\caption{\label{tab_final_list}Final software list}
\end{table}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.7]{figures/PrimaryLanguages.pdf}
\centering
\caption{\label{fig_language}Primary languages versus number of projects}
\end{figure}

@@

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\hline
\multirow{2}{*}{Software} & \multirow{2}{*}{Text files} & \multirow{2}{*}{Total lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ \hline
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ \hline
\end{tabular}
\caption{\label{tab_loc_per_file}Number of files and lines}
\end{table}

@@

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\hline
Software & Our ranking & Stars/yr & Watches/yr & Forks/yr \\ \hline
3D Slicer & 1 & 284 & 19 & 128 \\
OHIF Viewer & 3 & 277 & 19 & 224 \\
dwv & 23 & 124 & 12 & 51 \\
ImageJ & 2 & 84 & 9 & 30 \\
ParaView & 5 & 67 & 7 & 28 \\
Horos & 12 & 49 & 9 & 18 \\
Papaya & 17 & 45 & 5 & 20 \\
Fiji & 4 & 44 & 5 & 21 \\
DICOM Viewer & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 40 & 4 & 17 \\
Weasis & 6 & 36 & 5 & 19 \\
dicompyler & 26 & 35 & 5 & 14 \\
OsiriX Lite & 9 & 34 & 9 & 24 \\
MRIcroGL & 18 & 24 & 3 & 3 \\
GATE & 25 & 19 & 6 & 26 \\
Ginkgo CADx & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 8 & 18 & 5 & 7 \\
Drishti & 27 & 16 & 4 & 4 \\
Slice:Drop & 20 & 10 & 2 & 5 \\
ITK-SNAP & 15 & 9 & 1 & 4 \\
medInria & 7 & 7 & 3 & 6 \\
SMILI & 13 & 3 & 1 & 2 \\
MatrixUser & 28 & 2 & 0 & 0 \\
MicroView & 16 & 1 & 1 & 1 \\
Gwyddion & 11 & n/a & n/a & n/a \\
XMedCon & 19 & n/a & n/a & n/a \\
DicomBrower & 21 & n/a & n/a & n/a \\
AMIDE & 22 & n/a & n/a & n/a \\
3DimViewer & 24 & n/a & n/a & n/a \\ \hline
\end{tabular}
\caption{\label{tab_ranking_vs_GitHub}Software ranking versus GitHub metrics}
\end{table}
\endgroup

>>>>

Table \ref{tab_final_list} shows the 29 software packages that we measured,
along with summary data collected in the year 2020. We arrange the items in
descending order of LOC. We found the initial release dates (Rlsd) for most
projects and marked the two unknown dates with ``?''. We used the date of the
latest change to each code repository to decide the latest update. We found
funding information (Fnd) for only eight projects.  For the number of
contributors (NOC) we considered anyone who made at least one accepted commit as
a contributor. The NOC is not usually the same as the number of long-term
project members, since many projects received change requests and code from the
community.  With respect to the OS, 25 packages work on all three OSs: Windows
(W), macOS (M), and Linux (L). Although the usual approach to cross-platform
compatibility was to work natively on multiple OSes, five projects achieved
platform-independence via web applications. The full measurement data for all
packages is available in \citet{Dong2021-Data}.

\begin{table}[!ht]
\centering
\begin{tabular}{p{6cm}lllllllll}
\hline
\multirow{2}{*}{Software} & \multirow{2}{*}{Rlsd} & \multirow{2}{*}{Updated} & \multirow{2}{*}{Fnd} & \multirow{2}{*}{NOC} & \multirow{2}{*}{LOC} & \multicolumn{3}{c}{OS} & \multirow{2}{*}{Web} \\ \cline{7-9}
 &  &  &  &  &  & W & M & L &  \\ \hline
ParaView \citep{Ahrens2005} & 2002 & 2020-10 & X & 100 & 886326 & X & X & X & X \\
Gwyddion \citep{Nevcas2012} & 2004 & 2020-11 &  & 38 & 643427 & X & X & X &  \\
Horos \citep{horosproject2020} & ? & 2020-04 &  & 21 & 561617 &  & X &  &  \\
OsiriX Lite \citep{PixmeoSARL2019} & 2004 & 2019-11 &  & 9 & 544304 &  & X &  &  \\
3D Slicer \citep{Kikinis2014} & 1998 & 2020-08 & X & 100 & 501451 & X & X & X &  \\
Drishti \citep{Limaye2012} & 2012 & 2020-08 &  & 1 & 268168 & X & X & X &  \\
Ginkgo CADx \citep{Wollny2020} & 2010 & 2019-05 &  & 3 & 257144 & X & X & X &  \\
GATE \citep{Jan2004} & 2011 & 2020-10 &  & 45 & 207122 &  & X & X &  \\
3DimViewer \citep{TESCAN2020} & ? & 2020-03 & X & 3 & 178065 & X & X &  &  \\
medInria \citep{Fillard2012} & 2009 & 2020-11 &  & 21 & 148924 & X & X & X &  \\
BioImage Suite Web \citep{Papademetris2005} & 2018 & 2020-10 & X & 13 & 139699 &
X & X & X & X \\
Weasis \citep{Roduit2021} & 2010 & 2020-08 &  & 8 & 123272 & X & X & X &  \\
AMIDE \citep{Loening2017} & 2006 & 2017-01 &  & 4 & 102827 & X & X & X &  \\
XMedCon \citep{Nolf2003} & 2000 & 2020-08 &  & 2 & 96767 & X & X & X &  \\
ITK-SNAP \citep{Yushkevich2006} & 2006 & 2020-06 & X & 13 & 88530 & X & X & X &  \\
Papaya \citep{UTHSCSA2019} & 2012 & 2019-05 &  & 9 & 71831 & X & X & X &  \\
OHIF Viewer \citep{Ziegler2020} & 2015 & 2020-10 &  & 76 & 63951 & X & X & X & X \\
SMILI \citep{Chandra2018} & 2014 & 2020-06 &  & 9 & 62626 & X & X & X &  \\
INVESALIUS 3 \citep{Amorim2015} & 2009 & 2020-09 &  & 10 & 48605 & X & X & X &  \\
dwv \citep{Martelli2021} & 2012 & 2020-09 &  & 22 & 47815 & X & X & X & X \\
DICOM Viewer \citep{Afsar2021} & 2018 & 2020-04 & X & 5 & 30761 & X & X & X &  \\
MicroView \citep{ParallaxInnovations2020} & 2015 & 2020-08 &  & 2 & 27470 & X & X & X &  \\
MatrixUser \citep{Liu2016} & 2013 & 2018-07 &  & 1 & 23121 & X & X & X &  \\
Slice:Drop \citep{Haehn2013} & 2012 & 2020-04 &  & 3 & 19020 & X & X & X & X \\
dicompyler \citep{Panchal2010} & 2009 & 2020-01 &  & 2 & 15941 & X & X &  &  \\
Fiji \citep{Schindelin2012} & 2011 & 2020-08 & X & 55 & 10833 & X & X & X &  \\
ImageJ \citep{Rueden2017} & 1997 & 2020-08 & X & 18 & 9681 & X & X & X &  \\
MRIcroGL \citep{Rorden2021} & 2015 & 2020-08 &  & 2 & 8493 & X & X & X &  \\
DicomBrowser \citep{Archie2012} & 2012 & 2020-08 &  & 3 & 5505 & X & X & X &  \\ \hline
\end{tabular}
\caption{Final software list (sorted in descending order of the number of Lines
Of Code (LOC))}
\label{tab_final_list}
\end{table}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.5]{figures/PrimaryLanguages.pdf}
\centering
\caption{\label{fig_language}Primary languages versus number of projects}
\end{figure}

@@

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\hline
\multirow{2}{*}{Software} & \multirow{2}{*}{Text files} & \multirow{2}{*}{Total lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ \hline
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ \hline
\end{tabular}
\caption{Number of files and lines (entries are sorted in descending order of
their reusability scores)}
\label{tab_loc_per_file}
\end{table}

@@

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\hline
Software & Our ranking & Stars/yr & Watches/yr & Forks/yr \\ \hline
3D Slicer & 1 & 284 & 19 & 128 \\
OHIF Viewer & 3 & 277 & 19 & 224 \\
dwv & 23 & 124 & 12 & 51 \\
ImageJ & 2 & 84 & 9 & 30 \\
ParaView & 5 & 67 & 7 & 28 \\
Horos & 12 & 49 & 9 & 18 \\
Papaya & 17 & 45 & 5 & 20 \\
Fiji & 4 & 44 & 5 & 21 \\
DICOM Viewer & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 40 & 4 & 17 \\
Weasis & 6 & 36 & 5 & 19 \\
dicompyler & 26 & 35 & 5 & 14 \\
OsiriX Lite & 9 & 34 & 9 & 24 \\
MRIcroGL & 18 & 24 & 3 & 3 \\
GATE & 25 & 19 & 6 & 26 \\
Ginkgo CADx & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 8 & 18 & 5 & 7 \\
Drishti & 27 & 16 & 4 & 4 \\
Slice:Drop & 20 & 10 & 2 & 5 \\
ITK-SNAP & 15 & 9 & 1 & 4 \\
medInria & 7 & 7 & 3 & 6 \\
SMILI & 13 & 3 & 1 & 2 \\
MatrixUser & 28 & 2 & 0 & 0 \\
MicroView & 16 & 1 & 1 & 1 \\
Gwyddion & 11 & n/a & n/a & n/a \\
XMedCon & 19 & n/a & n/a & n/a \\
DicomBrower & 21 & n/a & n/a & n/a \\
AMIDE & 22 & n/a & n/a & n/a \\
3DimViewer & 24 & n/a & n/a & n/a \\ \hline
\end{tabular}
\caption{\label{tab_ranking_vs_GitHub}Software ranking versus GitHub metrics
(Sorted in descending order of GitHub popularity, as estimated by the number of
new stars per year)}
\end{table}
\endgroup
